Thesis Title

### 2022-05-23 chapters/chapter_2.tex
update thesis outline
mv tables in chapter 2 to folder

N. Couto, L. Schuele, E.C. Raangs, M. P. Machado, C. I. Mendes, T. F. Jesus, M. Chlebowicz,  S. Rosema, M. Ramirez, J. A. Carriço, I. B. Autenrieth, A. W. Friedrich, S. Peter and J. W. Rossen. Critical steps in clinical shotgun metagenomics for the concomitant detection and typing of microbial pathogens. Sci Rep 8, 13767 (2018).DOI: \url{https://doi.org/10.1038/s41598-018-31873-w}

@@

In this publication \ac{SMg} was applied to nine body fluid samples and one tissue sample from patients at the University Medical Center Groningen (UMCG) with varying degrees of contamination: one sample from peritoneal fluid, five from pus , two from synovial fluid of knees with prosthesis, one from sputum and one from a bone biopsy. The results of microbial identification through whole genome sequencing (\ac{WGS}) and \ac{SMg} were compared to standard culture-based microbiological methods.

In order to evaluate and compare the accuracy and reliability of the bioinformatics analyses in providing the closest results to culture and \ac{WGS} of any cultured isolates, three different bioinformatic pipelines (two commercially and one freely available) were used. Most pathogens identified by culture were also identified through metagenomics, but substantial differences were noted between the taxonomic classification tools.

@@

Nine body fluid samples and one tissue sample entering the Medical Microbiology laboratory were selected for metagenomics sequencing. 
These included one sample from peritoneal fluid, five from pus (3 abscesses and 2 empyema), two from synovial fluid of knees with prosthesis, one from sputum and one from a bone biopsy (Table \ref{tab:table_1}). 
All samples were stored at 4ºC for a variable period (2-10 days). 
The samples used for the present analyses were collected during routine diagnostics and infection prevention and control investigations. 
All procedures were carried out according to guidelines and regulations of University Medical Centre Groningen (UMCG) concerning the use of patient materials for the validation of clinical methods, which are in compliance with the guidelines of the Federation of Dutch Medical Scientific Societies (FDMSS).
Every patient entering the UMCG is informed that samples taken may be used for research and publication purposes, unless they indicate that they do not agree to it. 
This procedure has been approved by the Medical Ethical Committee of the UMCG. 
Informed consent was obtained from all individuals or their guardians prior to study participation. 
All samples were used after performing and completing a conventional microbiological diagnosis and were coded to protect patients’ confidentiality. 
All experiments were performed in accordance with the guidelines of the Declaration of Helsinki and the institutional regulations.

\begin{table}[]
\caption{Characteristics of the samples and mapping of trimmed reads against a human genome hg19 (\%) using CLC Genomics Workbench v10.0.1.}
\label{tab:ch2_table_1}
\resizebox{\textwidth}{!}{%
\begin{tabular}{@{}llllll@{}}
\toprule
\textbf{Sample} & \textbf{Sample type} & \textbf{DNA extraction method} & \textbf{Total number of reads} & \textbf{Mapped reads against hg19} & \textbf{Unmapped reads} \\ \midrule
Sample 1         & Peritoneal fluid & Ultra-Deep Microbiome Prep (Molzym) & 5892978 & 5,249,063 (89.2\%) & 632,951 (10.8\%)   \\
Sample 2         & Pus (abscess)    & Ultra-Deep Microbiome Prep (Molzym) & 9603346 & 7,828.746 (81.6\%) & 1,770,558 (18.4\%) \\
Sample 3         & Synovial fluid   & Ultra-Deep Microbiome Prep (Molzym) & 8615810 & 8,254,594 (95.9\%) & 355,200 (4.1\%)    \\
Sample 4         & Synovial fluid   & Ultra-Deep Microbiome Prep (Molzym) & 6078166 & 6,015,945 (99.0\%) & 61,099 (1.0\%)     \\
Sample 5         & Pus (abscess)    & Ultra-Deep Microbiome Prep (Molzym) & 8368930 & 309,588 (3.7\%)    & 8,052,272 (96.3\%) \\
Sample 6         & Pus (empyema)    & QIAamp DNA Microbiome Kit (Qiagen)  & 2912802 & 2,877,066 (98.8\%) & 34,506 (1.1\%)     \\
Sample 7         & Pus (empyema)    & QIAamp DNA Microbiome Kit (Qiagen)  & 1486700 & 922,932 (62.2\%)   & 561,772 (37.8\%)   \\
Sample 8         & Bone biopsy      & Micro-DXTM (Molzym)                 & 6534866 & 229,149 (3.5\%)    & 6,303,803 (96.5\%) \\
Sample 9         & Pus (abscess)    & Micro-DXTM (Molzym)                 & 6173132 & 6,081,612 (98.5\%) & 89,922 (1.5\%)     \\
Sample 10        & Sputum           & Micro-DXTM (Molzym)                 & 7596836 & 7,337,832 (96.7\%) & 235,520 (3.3\%)    \\
Negative control & Water            & QIAamp DNA Microbiome Kit (Qiagen)  & 1730738 & 1,706,861 (98.9\%) & 19,805 (1.2\%)     \\ \bottomrule
\end{tabular}%
}
\end{table}

@@

@@

\begin{table}[]
\caption{Microorganisms identified by conventional methods, \ac{WGS} and using shotgun metagenomics and the taxonomic classification methods in Unix.}
\label{tab:ch2_table2}
\resizebox{\textwidth}{!}{%
\begin{tabular}{@{}|l|l|l|l|lll|@{}}
\toprule
\multicolumn{1}{|c|}{\multirow{2}{*}{\textbf{Sample number}}} &
  \multicolumn{1}{c|}{\multirow{2}{*}{\textbf{Culture result (CFU)$^a$}}} &
  \multicolumn{1}{c|}{\multirow{2}{*}{\textbf{\begin{tabular}[c]{@{}c@{}}Conventional identification\\  (MALDI-TOF)\end{tabular}}}} &
  \multicolumn{1}{c|}{\multirow{2}{*}{\textbf{\ac{WGS}-based identification}}} &
  \multicolumn{3}{c|}{\textbf{Shotgun metagenomics}} \\ \cmidrule(l){5-7} 
\multicolumn{1}{|c|}{} &
  \multicolumn{1}{c|}{} &
  \multicolumn{1}{c|}{} &
  \multicolumn{1}{c|}{} &
  \multicolumn{1}{c|}{\textbf{Kraken$^b$}} &
  \multicolumn{1}{c|}{\textbf{MIDAS$^c$}} &
  \multicolumn{1}{c|}{\textbf{MetaPhlAn$^c$}} \\ \midrule
\textbf{1} &
  \begin{tabular}[c]{@{}l@{}}10$^3$\\  10$^3$\\  10\end{tabular} &
  \begin{tabular}[c]{@{}l@{}}\textit{E. faecium}\\  \textit{S. haemolyticus}\\  \textit{C. glabrata}\end{tabular} &
  \begin{tabular}[c]{@{}l@{}}\textit{E. faecium}\\  \textit{S. haemolyticus}\\  -\end{tabular} &
  \multicolumn{1}{l|}{\begin{tabular}[c]{@{}l@{}}\textit{E. faecium} (34.6\%)\\  \textit{S. haemolyticus} (10.1\%)\\  -\end{tabular}} &
  \multicolumn{1}{l|}{\begin{tabular}[c]{@{}l@{}}\textit{E. faecium} (62.0\%)\\  \textit{S. haemolyticus} (28.0\%)\\  -\end{tabular}} &
  \begin{tabular}[c]{@{}l@{}}\textit{E. faecium} (66.6\%)\\  \textit{S. haemolyticys} (27.7\%)\\  -\end{tabular} \\ \midrule
\textbf{2} &
  \begin{tabular}[c]{@{}l@{}}10$^3$\\  1\\  Not determined\end{tabular} &
  \begin{tabular}[c]{@{}l@{}}\textit{E. avium}\\  \textit{E. coli}\\  Anaerobes\end{tabular} &
  \begin{tabular}[c]{@{}l@{}}-\#\\  -\#\\  -\#\end{tabular} &
  \multicolumn{1}{l|}{\begin{tabular}[c]{@{}l@{}}Not identified$^*$\\  Not identified$^*$\\  Several species (29.5\%)\end{tabular}} &
  \multicolumn{1}{l|}{\begin{tabular}[c]{@{}l@{}}Not identified$^*$\\  Not identified$^*$\\  Several species (100.0\%)\end{tabular}} &
  \begin{tabular}[c]{@{}l@{}}Not identified$^*$\\  Not identified$^*$\\  Several species (100.0\%)\end{tabular} \\ \midrule
\textbf{3} &
  1 &
  \textit{S. epidermidis} &
  -$^\#$ &
  \multicolumn{1}{l|}{S. aureus (0.2\%)} &
  \multicolumn{1}{l|}{Not identified$^*$} &
  Not identified$^*$ \\ \midrule
\textbf{4} &
  10$^3$ &
  \textit{S. aureus} &
  \textit{S. aureus} &
  \multicolumn{1}{l|}{\textit{S. aureus} (0.73\%)} &
  \multicolumn{1}{l|}{\textit{S. aureus }(100\%)} &
  \textit{S. aureus} (100\%) \\ \midrule
\textbf{5} &
  \begin{tabular}[c]{@{}l@{}}$\geq$ 10$^5$\\  $\geq$ 10$^5$\\  10$^3$\\  10$^3$\\  Not determined\\  10\end{tabular} &
  \begin{tabular}[c]{@{}l@{}}\textit{E. coli}\\  \textit{K. oxytoca}\\  \textit{S. anginosus}\\  \textit{E. faecalis}\\  Anaerobes\\  \textit{C. albicans}\end{tabular} &
  \begin{tabular}[c]{@{}l@{}}\textit{E. coli}\\  \textit{K. oxytoca}\\  -$^\#$\\  \textit{E. faecalis}\\  -$^\#$\\  -$^\#$\end{tabular} &
  \multicolumn{1}{l|}{\begin{tabular}[c]{@{}l@{}}\textit{E. coli} (9.7\%)\\  \textit{K. oxytoca} (0.5\%)\\  \textit{S. anginosus} (0.07\%)\\  \textit{E. faecalis} (0.3\%)\\  Several species (12.7\%)\\  -\end{tabular}} &
  \multicolumn{1}{l|}{\begin{tabular}[c]{@{}l@{}}\textit{E. coli} (6.5\%)\\  \textit{K. oxytoca} (0.3\%)\\  \textit{S. anginosus} (0.01\%)\\  \textit{E. faecalis} (0.9\%)\\  Several species (96.7\%)\\  -\end{tabular}} &
  \begin{tabular}[c]{@{}l@{}}\textit{E. coli} (8.5\%)\\  \textit{K. oxytoca} (0.3\%)\\  \textit{Streptococcus spp.} (0.09\%)\\  \textit{E. faecalis} (0.7\%)\\  Several species (90.4\%)\\  -\end{tabular} \\ \midrule
\textbf{6} &
  10$^3$ &
  \textit{E. faecium} &
  \textit{E. faecium} &
  \multicolumn{1}{l|}{\textit{E. faecium} (0.77\%)} &
  \multicolumn{1}{l|}{Not identified$^*$} &
  Not identified$^*$ \\ \midrule
\textbf{7} &
  10$^2$ &
  \textit{S. aureus} &
  -$^\#$ &
  \multicolumn{1}{l|}{\textit{S. aureus} (82.9\%)} &
  \multicolumn{1}{l|}{\textit{S. aureus} (100\%)} &
  \textit{S. aureus} (100\%) \\ \midrule
\textbf{8} &
  10$^3$ &
  \textit{O. intermedium} &
  \textit{O. intermedium} &
  \multicolumn{1}{l|}{\textit{O. anthropi} (21.3\%)} &
  \multicolumn{1}{l|}{\textit{O. intermedium} (99.4\%)} &
  \textit{O. intermedium} (99.1\%) \\ \midrule
\textbf{9} &
  10$^3$ &
  \textit{S. aureus} &
  \textit{S. aureus} &
  \multicolumn{1}{l|}{\textit{S. aureus} (22.9\%)} &
  \multicolumn{1}{l|}{\textit{S. aureus} (100\%)} &
  \textit{S. aureus} (100\%) \\ \midrule
\textbf{10} &
  10$^3$ &
  \textit{S. marcescens} &
  -\# &
  \multicolumn{1}{l|}{\textit{S. marcescens} (64.7\%)} &
  \multicolumn{1}{l|}{\textit{S. marcescens} (99.1\%)} &
  \textit{S. marcescens} (100\%) \\ \bottomrule
\end{tabular}%
}
\tiny
\item $^a$The number of colonies of a given species was estimated from the number of colonies with the same morphology on the same plate 
\item $^b$The relative abundance is calculated using total number of reads as denominator
\item $^c$The relative abundance is calculated with the total number of classified reads as denominator
\item $^d$miniKraken database was used
\item $^\#$Although there was a laboratory identification, no isolates were available for \ac{WGS}
\item $^*$No reads matched that specific pathogen, not even at the genus level
\end{table}

\begin{table}[]
\caption{Microorganisms identified by conventional methods, \ac{WGS} and using shotgun metagenomics and the taxonomic classification methods in CLC Genomics Workbench.}
\label{tab:ch2_table3}
\resizebox{\textwidth}{!}{%
\begin{tabular}{@{}|l|l|l|l|ll|@{}}
\toprule
\multicolumn{1}{|c|}{\multirow{2}{*}{\textbf{Sample number}}} &
  \multicolumn{1}{c|}{\multirow{2}{*}{\textbf{Culture result (CFU)$^a$}}} &
  \multicolumn{1}{c|}{\multirow{2}{*}{\textbf{\begin{tabular}[c]{@{}c@{}}Conventional identification\\  (MALDI-TOF)\end{tabular}}}} &
  \multicolumn{1}{c|}{\multirow{2}{*}{\textbf{\ac{WGS}-based identification}}} &
  \multicolumn{2}{c|}{\textbf{Shotgun metagenomics}} \\ \cmidrule(l){5-6} 
\multicolumn{1}{|c|}{} &
  \multicolumn{1}{c|}{} &
  \multicolumn{1}{c|}{} &
  \multicolumn{1}{c|}{} &
  \multicolumn{1}{l|}{\textbf{Taxonomic Profiling (CLC)$^b$}} &
  \textbf{Best match with K-mer spectra (CLC)$^c$} \\ \midrule
\textbf{1} &
  \begin{tabular}[c]{@{}l@{}}103\\  10$^3$\\  10\end{tabular} &
  \begin{tabular}[c]{@{}l@{}}\textit{E. faecium}\\  \textit{S. haemolyticus}\\  \textit{C. glabrata}\end{tabular} &
  \begin{tabular}[c]{@{}l@{}}\textit{E. faecium}\\  \textit{S. haemolyticus}\\  -\end{tabular} &
  \multicolumn{1}{l|}{\begin{tabular}[c]{@{}l@{}}\textit{E. faecium} (71\%)\\  \textit{S. haemolyticus} (24\%)\\  \textit{C. glabrata} (100\%)\end{tabular}} &
  \begin{tabular}[c]{@{}l@{}}\textit{E. faecium} (41.4\%)\\  \textit{S. haemolyticus} (13.8\%)\\  \textit{C. glabrata} (0.5\%)\end{tabular} \\ \midrule
\textbf{2} &
  \begin{tabular}[c]{@{}l@{}}10$^3$\\  1\\  Not determined\end{tabular} &
  \begin{tabular}[c]{@{}l@{}}\textit{E. avium}\\  \textit{E. coli}\\  Anaerobes\end{tabular} &
  \begin{tabular}[c]{@{}l@{}}-$^\#$\\  -$^\#$\\  -$^\#$\end{tabular} &
  \multicolumn{1}{l|}{\begin{tabular}[c]{@{}l@{}}Not identified$^*$\\  Not identified$^*$\\  Several species (97\%)\end{tabular}} &
  \begin{tabular}[c]{@{}l@{}}Not identified$^*$\\  Not identified$^*$\\  Several species (13.2\%)\end{tabular} \\ \midrule
\textbf{3} &
  1 &
  \textit{S. epidermidis} &
  -\# &
  \multicolumn{1}{l|}{Not identified$^*$} &
  \textit{S. aureus} (4\%) \\ \midrule
\textbf{4} &
  10$^3$ &
  \textit{S. aureus} &
  \textit{S. aureus} &
  \multicolumn{1}{l|}{Not identified$^*$} &
  \textit{S. aureus} (9.7\%) \\ \midrule
\textbf{5} &
  \begin{tabular}[c]{@{}l@{}}$\geq$ 10$^5$\\  $\geq$ 10$^5$\\  10$^3$\\  10$^3$\\  Not determined\\  10\end{tabular} &
  \begin{tabular}[c]{@{}l@{}}\textit{E. coli}\\  \textit{K. oxytoca}\\  \textit{S. anginosus}\\  \textit{E. faecalis}\\  Anaerobes\\  \textit{C. albicans}\end{tabular} &
  \begin{tabular}[c]{@{}l@{}}\textit{E. coli}\\  \textit{K. oxytoca}\\  -$^\#$\\  E. faecalis\\  -$^\#$\\  -$^\#$\end{tabular} &
  \multicolumn{1}{l|}{\begin{tabular}[c]{@{}l@{}}\textit{E. coli} (25\%)\\  \textit{K. michiganensis} (0.3\%)\\  Not identified$^*$\\  \textit{E. faecalis} (2\%)\\  Several species (70.0\%)\\  Not identified$^*$\end{tabular}} &
  \begin{tabular}[c]{@{}l@{}}\textit{E. coli} (11.5\%)\\  Not identified$^*$\\  Not identified$^*$\\  \textit{E. faecalis} (0.6\%)\\  Not identified$^*$\\  \textit{C. albicans} (\textless{}0.05\%)\end{tabular} \\ \midrule
\textbf{6} &
  10$^3$ &
  \textit{E. faecium} &
  \textit{E. faecium} &
  \multicolumn{1}{l|}{Not identified$^*$} &
  \textit{E. faecium} (4.0\%) \\ \midrule
\textbf{7} &
  10$^2$ &
  \textit{S. aureus} &
  -$^\#$&
  \multicolumn{1}{l|}{\textit{S. aureus} (100\%)} &
  \textit{S. aureus} (95.5\%) \\ \midrule
\textbf{8} &
  10$^3$ &
  \textit{O. intermedium} &
  \textit{O. intermedium} &
  \multicolumn{1}{l|}{\textit{O. intermedium }(86.0\%)} &
  \textit{O. intermedium} (91.2\%) \\ \midrule
\textbf{9} &
  10$^3$ &
  \textit{S. aureus} &
  \textit{S. aureus} &
  \multicolumn{1}{l|}{\textit{S. aureus} (100\%)} &
  \textit{S. aureus} (81.2\%) \\ \midrule
\textbf{10} &
  10$^3$ &
  \textit{S. marcescens} &
  -$^\#$ &
  \multicolumn{1}{l|}{\textit{S. marscescens} (100\%)} &
  \textit{S. marcescens} (79.7\%) \\ \bottomrule
\end{tabular}%
}
\tiny
\item $^a$The number of colonies of a given species was estimated from the number of colonies with the same morphology on the same plate 
\item $^b$The relative abundance is calculated using total number of reads as denominator
\item $^c$The relative abundance is calculated with the total number of classified reads as denominator
\item $^\#$Although there was a laboratory identification, no isolates were available for \ac{WGS}
\item $^*$No reads matched that specific pathogen, not even at the genus level
\end{table}

\begin{table}[]
\caption{Microorganisms identified by conventional methods, \ac{WGS} and using shotgun metagenomics and the taxonomic classification methods in webpages (BaseSpace, Taxonomer and CosmosID).}
\label{tab:ch2_table4}
\resizebox{\textwidth}{!}{%
\begin{tabular}{@{}|l|l|l|l|lllll|@{}}
\toprule
\multirow{2}{*}{\textbf{Sample number}} &
  \multirow{2}{*}{\textbf{Culture result (CFU)$^a$}} &
  \multirow{2}{*}{\textbf{\begin{tabular}[c]{@{}l@{}}Conventional identification\\  (MALDI-TOF)\end{tabular}}} &
  \multirow{2}{*}{\textbf{\ac{WGS}-based identification}} &
  \multicolumn{5}{c|}{\textbf{Shotgun metagenomics}} \\ \cmidrule(l){5-9} 
 &
   &
   &
   &
  \multicolumn{1}{l|}{\textbf{\begin{tabular}[c]{@{}l@{}}Genius\\  (Basespace)$^c$\end{tabular}}} &
  \multicolumn{1}{l|}{\textbf{\begin{tabular}[c]{@{}l@{}}Kraken\\  (Basespace)$^c,d$\end{tabular}}} &
  \multicolumn{1}{l|}{\textbf{\begin{tabular}[c]{@{}l@{}}MetaPhlAn\\  (Basespace)$^c$\end{tabular}}} &
  \multicolumn{1}{l|}{\textbf{Taxonomer (Utah)$^b,e$}} &
  \textbf{Cosmos ID$^a$} \\ \midrule
\textbf{1} &
  \begin{tabular}[c]{@{}l@{}}10$^3$\\  10$^3$\\  10\end{tabular} &
  \begin{tabular}[c]{@{}l@{}}\textit{E. faecium}\\  \textit{S. haemolyticus}\\  \textit{C. glabrata}\end{tabular} &
  \begin{tabular}[c]{@{}l@{}}\textit{E. faecium}\\  \textit{S. haemolyticus}\\  -\end{tabular} &
  \multicolumn{1}{l|}{\begin{tabular}[c]{@{}l@{}}\textit{E. faecium} (14.4\%)\\  \textit{S. haemolyticus} (55.8\%)\\  -\end{tabular}} &
  \multicolumn{1}{l|}{\begin{tabular}[c]{@{}l@{}}\textit{E. faecium} (25.0\%)\\  \textit{S. haemolyticus} (20.1\%)\\  -\end{tabular}} &
  \multicolumn{1}{l|}{\begin{tabular}[c]{@{}l@{}}\textit{E. faecium} (65.1\%)\\  \textit{S. haemolyticys} (30.4\%)\\  -\end{tabular}} &
  \multicolumn{1}{l|}{\begin{tabular}[c]{@{}l@{}}\textit{E. faecium} (22.9\%)\\  \textit{S. haemolyticus} (20.1\%)\\  Not identified$^*$\end{tabular}} &
  \begin{tabular}[c]{@{}l@{}}\textit{E. faecium} (50.3\%)\\  \textit{S. haemolyticus} (22.1\%)\\  \textit{C. glabrata} (88.6\%)\end{tabular} \\ \midrule
\textbf{2} &
  \begin{tabular}[c]{@{}l@{}}103\\  1\\  Not determined\end{tabular} &
  \begin{tabular}[c]{@{}l@{}}\textit{E. avium}\\  \textit{E. coli}\\  Anaerobes\end{tabular} &
  \begin{tabular}[c]{@{}l@{}}-$^\#$\\  -$^\#$\\  -$^\#$\end{tabular} &
  \multicolumn{1}{l|}{\begin{tabular}[c]{@{}l@{}}Not identified$^*$\\  Not identified$^*$\\  Several species (94.0\%)\end{tabular}} &
  \multicolumn{1}{l|}{\begin{tabular}[c]{@{}l@{}}Not identified$^*$\\  Not identified$^*$\\  Several species (27.0\%)\end{tabular}} &
  \multicolumn{1}{l|}{\begin{tabular}[c]{@{}l@{}}Not identified$^*$\\  Not identified$^*$\\  Several species (54.2\%)\end{tabular}} &
  \multicolumn{1}{l|}{\begin{tabular}[c]{@{}l@{}}Not identified$^*$\\  Not identified$^*$\\  Several species (14.2\%)\end{tabular}} &
  \begin{tabular}[c]{@{}l@{}}Not identified$^*$\\  Not identified$^*$\\  Several species (100\%)\end{tabular} \\ \midrule
\textbf{3} &
  1 &
  \textit{S. epidermidis} &
  -\# &
  \multicolumn{1}{l|}{\textit{S. aureus} (100\%)} &
  \multicolumn{1}{l|}{\textit{S. aureus} (0.1\%)} &
  \multicolumn{1}{l|}{Not identified$^*$} &
  \multicolumn{1}{l|}{\textit{S. pseudintermedius} (3.4\%)} &
  Not identified$^*$ \\ \midrule
\textbf{4} &
  10$^3$ &
  \textit{S. aureus} &
  \textit{S. aureus} &
  \multicolumn{1}{l|}{\textit{S. aureus} (100\%)} &
  \multicolumn{1}{l|}{\textit{S. aureus} (0.3\%)} &
  \multicolumn{1}{l|}{\textit{S. aureus} (100\%)} &
  \multicolumn{1}{l|}{\textit{S. aureus} (8.3\%)} &
  \textit{S. aureus} (100\%) \\ \midrule
\textbf{5} &
  \begin{tabular}[c]{@{}l@{}}$\geq$ 10$^5$\\  $\geq$ 10$^5$\\  10$^3$\\  10$^3$\\  Not determined\\  10\end{tabular} &
  \begin{tabular}[c]{@{}l@{}}\textit{E. coli}\\  \textit{K. oxytoca}\\  \textit{S. anginosus}\\  \textit{E. faecalis}\\  Anaerobes\\  \textit{C. albicans}\end{tabular} &
  \begin{tabular}[c]{@{}l@{}}\textit{E. coli}\\  \textit{K. oxytoca}\\  -$^\#$\\  \textit{E. faecalis}\\  -$^\#$\\  -$^\#$\end{tabular} &
  \multicolumn{1}{l|}{\begin{tabular}[c]{@{}l@{}}\textit{E. coli} (0.4\%)\\  Not identified$^*$\\  \textit{S. anginosus} (0.03\%)\\  \textit{E. faecalis} (0.8\%)\\  Several species (45.0\%)\\  -\end{tabular}} &
  \multicolumn{1}{l|}{\begin{tabular}[c]{@{}l@{}}\textit{E. coli} (10.2\%)\\  \textit{K. oxytoca} (0.5\%)\\  \textit{S. anginosus} (0.4\%)\\  \textit{E. faecalis} (0.3\%)\\  Several species (8.0\%)\\  -\end{tabular}} &
  \multicolumn{1}{l|}{\begin{tabular}[c]{@{}l@{}}\textit{E. coli} (7.0\%)\\  \textit{K. pneumoniae} (0.01\%)\\  \textit{S. anginosus} (0.3\%)\\  \textit{E. faecalis} (0.7\%)\\  Several species (89.1\%)\\  -\end{tabular}} &
  \multicolumn{1}{l|}{\begin{tabular}[c]{@{}l@{}}\textit{E. coli} (3.6\%)\\  \textit{K. michiganensis} (0.1\%)\\  \textit{S. anginosus} (0.1\%)\\  \textit{E. faecalis} (0.1\%)\\  Several species (60.3\%)\\  -\end{tabular}} &
  \begin{tabular}[c]{@{}l@{}}\textit{E. coli} (7.6\%)\\  \textit{K. oxytoca} (1.7\%)\\  \textit{S. anginosus} (0.09\%)\\  \textit{E. faecalis} (3.7\%)\\  Several species (86.2\%)\\  Not identified$^*$\end{tabular} \\ \midrule
\textbf{6} &
  10$^3$ &
  \textit{E. faecium} &
  \textit{E. faecium} &
  \multicolumn{1}{l|}{\textit{E. faecium} (4.2\%)} &
  \multicolumn{1}{l|}{\textit{E. faecium} (14.8\%)} &
  \multicolumn{1}{l|}{\textit{E. faecium} (5.5\%)} &
  \multicolumn{1}{l|}{\textit{E. faecium} (1.4\%)} &
  \textit{E. faecium} (4.1\%) \\ \midrule
\textbf{7} &
  10$^2$ &
  \textit{S. aureus} &
  -\# &
  \multicolumn{1}{l|}{\textit{S. aureus} (100\%)} &
  \multicolumn{1}{l|}{\textit{S. aureus} (93.8\%)} &
  \multicolumn{1}{l|}{\textit{S. aureus} (100\%)} &
  \multicolumn{1}{l|}{\textit{S. aureus} (14.2\%)} &
  \textit{S. aureus} (100\%) \\ \midrule
\textbf{8} &
  10$^3$ &
  \textit{O. intermedium} &
  \textit{O. intermedium} &
  \multicolumn{1}{l|}{\textit{O. intermedium} (100\%)} &
  \multicolumn{1}{l|}{\textit{O. nthropic} (88.9\%)} &
  \multicolumn{1}{l|}{\textit{O. intermedium} (99.8\%)} &
  \multicolumn{1}{l|}{\textit{O. intermedium} (13.1\%)} &
  \textit{O. intermedium} (49.5\%) \\ \midrule
\textbf{9} &
  10$^3$ &
  \textit{S. aureus} &
  \textit{S. aureus} &
  \multicolumn{1}{l|}{\textit{S. aureus} (100\%)} &
  \multicolumn{1}{l|}{\textit{S. aureus} (99.5\%)} &
  \multicolumn{1}{l|}{\textit{S. aureus} (100\%)} &
  \multicolumn{1}{l|}{\textit{S. aureus} (12.7\%)} &
  \textit{S. aureus} (100\%) \\ \midrule
\textbf{10} &
  10$^3$ &
  \textit{S. marcescens} &
  -$^\#$ &
  \multicolumn{1}{l|}{\textit{S. marcescens} (32.5\%)} &
  \multicolumn{1}{l|}{\textit{S. marcescens} (94.8\%)} &
  \multicolumn{1}{l|}{\textit{Serratia} spp. (100\%)} &
  \multicolumn{1}{l|}{\textit{S. marcescens} (1.4\%)} &
  S\textit{. marscescens} (38.4\%) \\ \bottomrule
\end{tabular}%
}
\tiny
\item $^a$The number of colonies of a given species was estimated from the number of colonies with the same morphology on the same plate 
\item $^b$The relative abundance is calculated using total number of reads as denominator
\item $^c$The relative abundance is calculated with the total number of classified reads as denominator
\item $^d$miniKraken database was used
\itel $^e$Full Analysis mode was used 
\item $^\#$Although there was a laboratory identification, no isolates were available for \ac{WGS}
\item $^*$No reads matched that specific pathogen, not even at the genus level
\end{table}

@@

\begin{table}[]
\caption{Performance of the different taxonomic classification methods for each sample. Sensitivity and positive predictive value were calculated using culture/MALDI-TOF as standards}
\label{tab:ch2_table5}
\resizebox{\textwidth}{!}{%
\begin{tabular}{@{}lllllll@{}}
\toprule
\textbf{Method} &
  \textbf{Total number of bacteria identified$^a$} &
  \textbf{True positives$^a$} &
  \textbf{False positives} &
  \textbf{False negatives} &
  \textbf{Sensitivity (\%)} &
  \textbf{PPV (\%)} \\ \midrule
Culture/MALDI-TOF                                         & 9    & 9 & 0    & 0 & 1    & 1    \\
MetaPhlAn (BaseSpace)                                     & 16   & 7 & 9    & 2 & 0.78 & 0.44 \\
Genius (BaseSpace)                                        & 35   & 8 & 27   & 1 & 0.89 & 0.23 \\
Kraken (BaseSpace)                                        & 959  & 7 & 952  & 2 & 0.78 & 0.01 \\
Taxonomer (Full Analysis)                                 & 4649 & 8 & 4641 & 1 & 0.89 & 0    \\
CosmosID                                                  & 35   & 8 & 27   & 1 & 0.89 & 0.23 \\
Taxonomic Profiling (CLC Genomics Workbench v10.0.1)      & 17   & 6 & 11   & 3 & 0.67 & 0.35 \\
Best match K-mer spectra (CLC Genomics Workbench v10.0.1) & 12   & 8 & 4    & 1 & 0.89 & 0.67 \\
Kraken (Unix)                                             & 198  & 7 & 191  & 2 & 0.78 & 0.04 \\
MetaPhlAn2 (Unix)                                         & 15   & 7 & 6    & 4 & 0.75 & 0.75 \\
MIDAS (Unix)                                              & 34   & 7 & 26   & 2 & 0.88 & 0.5 \\ \bottomrule
\end{tabular}%
}
\tiny
\item $^a$Excluding the samples with non-identified anaerobic bacteria (Samples 2 and 5)
\item Abbreviations: PPV – positive predictive value
\end{table}

@@

@@

\begin{longtable}{ll}
\caption{Supplementary table 3.}
\label{tab:ch2_suptable3}\\
\multicolumn{2}{l}{Illumina}                                                                  \\
\endfirsthead
%
\endhead
%
Discard sequence names                       & FALSE                                          \\
Discard quality scores                       & FALSE                                          \\
Selected files                               &                                                \\
Paired-end reads                             & TRUE                                           \\
Read Orientation                             & Forward Reverse                                \\
minimum distance                             & 1                                              \\
maximum distance                             & 1000                                           \\
Remove failed reads                          & TRUE                                           \\
Quality score                                & NCBI/Sanger or Illumina Pipeline 1.8 and later \\
MiSeq de-multiplexing                        & FALSE                                          \\
Illumina trim                                & FALSE                                          \\
                                             &                                                \\
Trim Reads                                   &                                                \\
Quality trim                                 & TRUE                                           \\
Quality limit                                & 0.05                                           \\
Ambiguous trim                               & TRUE                                           \\
Ambiguous limit                              & 2                                              \\
Trim adapter list                            & Illumina adapters                              \\
Use colorspace                               & FALSE                                          \\
Remove 5' terminal nucleotides               & FALSE                                          \\
Number of 5' terminal nucleotides            & 1                                              \\

>>>>

N. Couto, L. Schuele, E.C. Raangs, M. P. Machado, C. I. Mendes, T. F. Jesus, M. Chlebowicz,  S. Rosema, M. Ramirez, J. A. Carriço, I. B. Autenrieth, A. W. Friedrich, S. Peter and J. W. Rossen. Critical steps in clinical shotgun metagenomics for the concomitant detection and typing of microbial pathogens. Sci Rep 8, 13767 (2018). DOI: \url{https://doi.org/10.1038/s41598-018-31873-w}

@@

In this publication \ac{SMg} was applied to nine body fluid samples and one tissue sample from patients at the University Medical Center Groningen (UMCG) with varying degrees of contamination: one sample from peritoneal fluid, five from pus , two from synovial fluid of knees with prosthesis, one from sputum and one from a bone biopsy. The results of microbial identification through whole genome sequencing (\ac{WGS}) and \ac{SMg} were compared to standard culture-based microbiological methods. 
In order to evaluate and compare the accuracy and reliability of the bioinformatics analyses in providing the closest results to culture and \ac{WGS} of any cultured isolates, three different bioinformatic pipelines (two commercially and one freely available) were used. Most pathogens identified by culture were also identified through metagenomics, but substantial differences were noted between the taxonomic classification tools.

@@

Nine body fluid samples and one tissue sample entering the Medical Microbiology laboratory were selected for metagenomics sequencing. 
These included one sample from peritoneal fluid, five from pus (3 abscesses and 2 empyema), two from synovial fluid of knees with prosthesis, one from sputum and one from a bone biopsy (Table \ref{tab:ch2_table_1}). 
All samples were stored at 4ºC for a variable period (2-10 days). 
The samples used for the present analyses were collected during routine diagnostics and infection prevention and control investigations. 
All procedures were carried out according to guidelines and regulations of University Medical Centre Groningen (UMCG) concerning the use of patient materials for the validation of clinical methods, which are in compliance with the guidelines of the Federation of Dutch Medical Scientific Societies (FDMSS).
Every patient entering the UMCG is informed that samples taken may be used for research and publication purposes, unless they indicate that they do not agree to it. 
This procedure has been approved by the Medical Ethical Committee of the UMCG. 
Informed consent was obtained from all individuals or their guardians prior to study participation. 
All samples were used after performing and completing a conventional microbiological diagnosis and were coded to protect patients’ confidentiality. 
All experiments were performed in accordance with the guidelines of the Declaration of Helsinki and the institutional regulations.

@@

\begin{landscape}
\input{tables/chapter_2/table1}
\end{landscape}

@@

\begin{landscape}
\input{tables/chapter_2/table2}
\end{landscape}

\begin{landscape}
\input{tables/chapter_2/table3}
\end{landscape}

\begin{landscape}
\input{tables/chapter_2/table4}
\end{landscape}

@@

\begin{landscape}
\input{tables/chapter_2/table5}
\end{landscape}

@@

\begin{landscape}
\input{tables/chapter_2/table6}
\end{landscape}

\begin{landscape}
\input{tables/chapter_2/table7}
\end{landscape}

@@

\input{tables/chapter_2/suptable3}
